%!TEX root = ../thesis-main.tex
\begin{acknowledgements} % this is optional
I would like to acknowledge and express my heartfelt appreciation to the wonderful people who have supported and encouraged me throughout the journey of completing my Master's thesis.\newline

Firstly, I would like to thank my family, including my mother Alessandra, my father Manuel, and my brother Enea, for their unwavering love, patience, and belief in my abilities. Their constant encouragement and support have been invaluable to me.\newline

I would also like to express my gratitude to my loving girlfriend Giorgia, who has been a constant source of support and love throughout my studies. She also deserves special thanks for her excellent English teaching skills, which were particularly helpful in correcting small details in this thesis.\newline

I am incredibly grateful to all of my friends in the Bewolla gang for always being there for me during the ups and downs. I am grateful for all the fun moments we shared together, from creating new words for the "Vocabolangelo" project to playing "The Game". These playful moments provided a much-needed break and helped me maintain my mental and emotional well-being. My heartfelt thanks also goes to my friends and colleagues at the university for their inspiration and motivation to always do better. The countless hours of study, project work, and collaboration have been critical to my personal and academic growth.\newline

Finally, I would like to extend my deepest gratitude to my supervisors, Danilo Pianini and Gianluca Aguzzi, for their guidance, patience, and support throughout this journey, and for giving me the opportunity to work on a very complex and
sophisticated project.\newline

Without the support of these incredible people, this thesis would not have been possible. I am truly grateful for their contributions, and I look forward to carrying their wisdom and support with me into the next chapter of my life.
\end{acknowledgements}
